## Functions and Conditionals

### Conditionals

The basic syntax for an if statement goes like this:

```{r}
if (logical statement) {
  do something
} else {
  do something else
}
```

The logical statements should boil down to a boolean; for example:

```{r}
a <- 15
if (a>10){
  print("Huge")
}
```

We can also add else if statements in the middle if we need more options.

```{r}
a <- 15
if (a<5){
  print(a)
} else {
  print("Its more than 5")
}
if (a<5){
  print("small")
} else if (a> 5 & a<= 10) {
  print("medium")
} else if (a > 10 | a == 7){
  print("Large")
} else {
  print("Huge")
}
```

Neat! We can also nest conditionals just like we do in Python.

```{r}
a<-5
if(a/2 != 0){
  if(a==5){
    print("a is 5")
  }
}
    
```

Neat! Now, let's look at another core part of R programming - Loops!

### Loops

```{r}
for (value in 1:5){
  #Paste is a concat function
  print(paste(value, "is greater than ", value -1))
}
```

We can also do nifty stuff with our vectors, such as apply changes to specific indices of our data.

```{r}
a_vector <- c(1,1,1,1,1,1,1)
for (i in c(1,3,6)){
  a_vector[i] <-6
}
a_vector
```

We can loop over more than just lists:

```{r}
mat <- matrix(1, nrow <- 2, ncol <- 2)
mat

for (row in 1:2){
  for (col in 1:2){
    mat[row,col] <- 0
    print(mat)
  }
}
```

Like Python, we have while loops in R:

```{r}
b <- 0
while (b<10){
  # increase b by 2
  b <- b+2
  print(b)
}
```

## Functions

When should we use functions?

```{r}
our_function <- function(optional_params){
  # do some stuff
}
test_temp <- c(16.55,26.9,32.5,8.5,28.0,28.5,20.9,21.1,21.0,18.7)

#Then we want a function that converst from C to F
c_to_f <- function(dat){
  temp_fahr <- dat*(9/5)+32
  return(temp_fahr)
  
}
print(c_to_f(26.1))
```

An example:

```{r}
print(c_to_f(26.1))
```

This is all cool.. but it didn't do anything. We need to call the function!

```{r}
data_fahr <- c_to_f(test_temp)
print(data_fahr)
```

Notice that in the background here, we are looping over each item in the array, but we never write a for loop! Keep this in mind, since when you pass in a large data set, the changes made by functions aren't quite immediate - the computer does have to go over each item in your data set, and change them one at a time.

A shortcut of sorts is to use the `apply()` function, which operates very similarly to Python's `map()`. Note that `apply` syntax goes as follows: `apply(X, MARGIN, FUN)` where X is the data set to use, MARGIN is specification for rows or columns, and FUN is the function to apply.

```{r}
test_mat <- matrix(c(1,1,1,2,2,2), nrow <- 2)
#test_mat
#using the functioon apply, we pass are matrix with specificed number or rows or columns and we sum across them using the sum function. 
apply(test_mat, 2, sum)
```

It is common to use `mean`, `median`, `sum`, `min`, or `max` with `apply`, but we can also use our user-defined functions as well. Note that apply() only works with matrices.

```{r}
test_temp_mat <- matrix(c(16.55,26.9,32.5,8.5,28.0,28.5,20.9,21.1,21.0,18.7), nrow = 2)
test_temp_mat

apply(test_temp_mat, 1, c_to_f)
```

### Homework

1.  Do 5 CodeWars problems in R. Share your solutions here:

    ```{r}
  #Lario and Muigi Pipe Problem
    pipe_fix <- function(nums){
  my_vector = c()
  last_element <- length(nums)
  for (index in nums[1]:nums[last_element]){
    my_vector = c(my_vector,index)
  }
  return(my_vector)
}
#Switcheroo
    switcheroo <- function(string){
 c = ""
  for (ele in strsplit(string, NULL)[[1]]) {
    if (ele == "a"){
      c<-paste(c,'b', sep = "")
    }
    else if( ele == "b"){
      c<-paste(c,'a', sep = "")
    }
     else{
     c<- paste(c,"c",sep = "")
       }
      }
  return (c)
    }
#DNA to RNA conversion
dna_to_rna <- function(dna){
   c <- ""
  for (ele in strsplit(dna, NULL)[[1]]) {
    if (ele == "T") {
      c <- paste(c, 'U', sep = "")
     } else {
      c <- paste(c, ele, sep = "")
    }
  } 
  return(c)
}
#how much do I love you
how_much_i_love_you <- function(nbpetals){
 count = 0
 count <- nbpetals%%6
  if (count == 1){
    return ("I love you")
  }
  else if (count == 2){
    return ("a little")
  }
  else if (count == 3){
    return ("a lot")
  }
  else if (count == 4){
    return ("passionately")
  }
  else if (count == 5){
    return ("madly")
  }
  else if (count == 0){
    return ("not at all")
  }
}
#Find the position!
position <- function(alphabet){
  alphabet <- tolower(alphabet)
 if (alphabet == "a"){
   return("Position of alphabet: 1")
 }
  if (alphabet == "b"){
   return("Position of alphabet: 2")
 }
  if (alphabet == "c"){
   return("Position of alphabet: 3")
 }
  if (alphabet == "d"){
   return("Position of alphabet: 4")
 }
  if (alphabet == "e"){
   return("Position of alphabet: 5")
 }
  if (alphabet == "f"){
   return("Position of alphabet: 6")
 }
  if (alphabet == "g"){
   return("Position of alphabet: 7")
 }
  if (alphabet == "h"){
   return("Position of alphabet: 8")
 }
  if (alphabet == "i"){
   return("Position of alphabet: 9")
 }
  if (alphabet == "j"){
   return("Position of alphabet: 10")
 }
  if (alphabet == "k"){
   return("Position of alphabet: 11")
}
  if (alphabet == "l"){
   return("Position of alphabet: 12")
}
  if (alphabet == "m"){
   return("Position of alphabet: 13")
}
  if (alphabet == "n"){
   return("Position of alphabet: 14")
}
  if (alphabet == "o"){
   return("Position of alphabet: 15")
}
  if (alphabet == "p"){
   return("Position of alphabet: 16")
}
  if (alphabet == "q"){
   return("Position of alphabet: 17")
}
  if (alphabet == "r"){
   return("Position of alphabet: 18")
}
  if (alphabet == "s"){
   return("Position of alphabet: 19")
}
  if (alphabet == "t"){
   return("Position of alphabet: 20")
}
  if (alphabet == "u"){
   return("Position of alphabet: 21")
}
  if (alphabet == "v"){
   return("Position of alphabet: 22")
}
  if (alphabet == "w"){
   return("Position of alphabet: 23")
}
  if (alphabet == "x"){
   return("Position of alphabet: 24")
}
  if (alphabet == "y"){
   return("Position of alphabet: 25")
}
  if (alphabet == "z"){
   return("Position of alphabet: 26")
}
#is a,e,i,o,u 
   is_vow <- function(a){
  for (i in 1:length(a)) {
    if (a[i] == 65 || a[i] == 97) {
      a[i] <- "a"
    }
    if (a[i] == 69 || a[i] == 101) {
      a[i] <- "e"
    }
    if (a[i] == 73 || a[i] == 105) {
      a[i] <- "i"
    }
    if (a[i] == 80 || a[i] == 111) {
      a[i] <- "o"
    }
    if (a[i] == 85 || a[i] == 117) {
      a[i] <- "u"
    }
  }
   return (a)
  }
  
    ```

2.  Create a function that checks an input of type `int` and returns fizz if the number is a multiple of 3, buzz if the number is a multiple of 5, and fizzbuzz if the number is a multiple of 3 and 5.

    ```{r}
  fizzbuzz <- function(int){
   if (int %% 1 != 0) {
    return("Not an integer")
  }
  if (int%%3==0 && int%%5==0){
   return("fizzbuzz")
  } else if (int%%5==0){
    return("buzz")
  } else if (int%%3==0){
    return("fizz")
  } else {
    return("None")
  }
  }


    ```

3.  **Mario Challenge:** In the classic Super Mario games, Mario ends each level by ascending a flight of stairs and leaping off the top to arrive at the castle at the end. (See [here](https://youtu.be/-avspZlbOWU?t=55) for details). We can imagine that each of the bricks of the stairs is a character X. Create a function that prints out a set of stairs. For example:

    If the function is called marioStairs() and we pass in the value 4, we would expect to see stairs with a length of 4 and a height of 4:

    X

    XX

    XXX

    XXXX

    ```{r}
    # You may feel more comfortable figuring this out in Python first, then translating into R later.
    ```
