## **R Libraries: ggplot2**

The R library `ggplot2` is a super useful tool we can use to create visualizations and explore our data.

```{bash}
install.packages("tidyverse")
library("tidyverse")
```

More things we can do to get more information on R and what we're doing:

```{bash}
help()
```
library("tidyverse")

We'll need a few more packages for our exploration usage, too. We can install them all at once by using the `c()` function, which creates a list. We'll open these packages with the `library()` function later.

```{bash}
install.packages(c("nycflights13", "gapminder", "Lahman"))
```

Built in to our tidyverse are several data sets that we can play around with. In the below code block, we create a variable `myplot` and save the `mpg` database to it.

```{r}
myplot <- mpg
myplot

```

**Something you should be aware of and have open (perhaps on a second monitor, or you can print them out) is the ggplot2 cheat sheet!** <https://github.com/rstudio/cheatsheets/blob/main/data-visualization-2.1.pdf>

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
  
```

```{r}
mtcars
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y = hwy, color = class))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y = hwy, color = displ < 5))

```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y = hwy, color = rgb(0.7,1,0.3))) +
  facet_wrap(~class, nrow = 2)


```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y = hwy, color = drv))+
  facet_grid(drv ~ .)
```

```{r}
#Use gg.plot cheat sheet
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))  +
  geom_point(mapping = aes(x= displ, y = hwy, color = drv))

ggplot(data = mpg, mapping = aes(x = displ, y= hwy, linetype = drv))+geom_point(mapping = aes(color = class))+
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))

```

Whoa, where'd count come from? **Count is not a variable or attribute of the variable diamonds!** Count comes automatically with a lot of graphs as an attribute. You can verify what I'm saying with `?diamonds` in the console. It shows up due to nature of discrete data sets

Any attribute of a data set that is algorithmically calculated is called a *stat*, which is short for a statistical transformation. Many of the **geom** functions have stats built in, and many stats display geoms. For example, the above code block used a geom, but this one uses a stat, and results in an identical chart:

```{r}
ggplot(data = diamonds) +
  stat_count(mapping = aes(x = cut))
```

Let's add color here!

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
  
```

What if we add to a different variable than cut?

```{r}
pie = ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity),
           position = "fill"
           ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

pie + coord_polar()
```

**For your consideration...** Is the above a good graph?

```{r}
#alt- becomes <- (option- on mac)

pie = ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity),
           position = "dodge"
           ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

pie + coord_polar()
```

## Homework

```{r}
# Find a builtin data set for students to explore; use their cheat sheet from stats to label continuous and discrete data; Analyze that data set, create graphs, find a hypothesis
library(ggplot2)
library(nycflights13)
#View(nycflights13::flights)
#View(Lahman::Batting)
#View(gapminder::gapminder)
#Generate 3 graphs to solve this homework 

#ggplot(data = nycflights13::flights) +
# geom_point(mapping = aes(x = dep_delay, y = sched_dep_time))


#Filter the data to exclude departure delays below 60
filtered_data <- flights %>%
  filter(dep_delay >= 60)

#ggplot(data = filtered_data, aes(x = dep_delay)) +
# geom_point(mapping = aes(x = dep_delay, y = sched_dep_time))


# Create a count plot (bar chart) of departure delays to show the frequency of long delays are less common
ggplot(data = filtered_data, aes(x = dep_delay)) +
  geom_bar() +
  coord_cartesian(xlim = c(60, 1000))

#Displaying a graph of frequency of sched_dep_time to show the most common usage of flights is in the morning,
#Noticing from the graph there is not a second hump of travel implying it does not follow rush our rules
#of having multiple busy times
ggplot(data = flights, aes(x = sched_dep_time)) +
  geom_bar() +
  labs(x = "Scheduled Time", y = "Count") +  # Customize axis labels
  ggtitle("Count of Flights by Scheduled Time") 

#ggplot(data = gapminder::gapminder) +
# geom_point(mapping = aes(x = lifeExp, y = gdpPercap))+
#   geom_smooth()

#Only using the most recent year of data to display life Expectancy relation to GDP
filtered_data <- gapminder::gapminder %>%
  filter(year == 2007)

ggplot(data = filtered_data, aes(x = lifeExp, y = gdpPercap)) +
  geom_point() +  # Add points
  geom_smooth(aes(group = 1))

#Only using the oldest data to show there is a significant change in the curve of relationship implying advances made in the modern era
filtered_data <- gapminder::gapminder %>%
  filter(year == 1952)

ggplot(data = filtered_data, aes(x = lifeExp, y = gdpPercap)) +
  geom_point() +  # Add points
  geom_smooth(aes(group = 1))
```

